---
title: "BFI"
author: "Karyn Huang"
date: "1/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE }
library(psych)
library(lavaan)
library(semTools)
```
## Processing the data 
```{r}
# select variables relevant to this study
load("/Users/mhhuang/OneDrive - UBC/BFI Study/Analysis/bfi.cleaned.data.participant.v1.RData")
data <- bfi.cleaned.data.participant.v1[,c(1:61,122:129)]

# filter data basing on condition 
likert_data <- subset(data, data$self_condition == "O") #“O” refers to the Original Likert version
expanded_data <- subset(data, data$self_condition == "E") #“E” refers to the Expanded version 
```


## Building the Models

### Extroversion 
#### Model Syntax
Model 0 has no method factor.
```{r}
Model0_E <- '
Sociability  =~ self_BFI_1 + self_BFI_16 + self_BFI_31 + self_BFI_46
Assertiveness  =~ self_BFI_6 + self_BFI_21 + self_BFI_36 + self_BFI_51
Energy_Level  =~ self_BFI_11 + self_BFI_26 + self_BFI_41 + self_BFI_56

#Extroversion =~ Sociability + Assertiveness + Energy_Level
'
```
Model 1 has an Acquiescence factor to capture the residuals variance of all items. 
```{r}
Model1_E <- '
Sociability  =~ self_BFI_1 + self_BFI_16 + self_BFI_31 + self_BFI_46
Assertiveness  =~ self_BFI_6 + self_BFI_21 + self_BFI_36 + self_BFI_51
Energy_Level  =~ self_BFI_11 + self_BFI_26 + self_BFI_41 + self_BFI_56

#Extroversion =~ Sociability + Assertiveness + Energy_Level

Sociability ~~ 0*Acquiescence
Energy_Level ~~ 0*Acquiescence
Assertiveness ~~ 0*Acquiescence

Acquiescence =~ 1*self_BFI_1 + (-1)*self_BFI_16 + (-1)*self_BFI_31 + 1*self_BFI_46 + 
1*self_BFI_6 + 1*self_BFI_21 + (-1)*self_BFI_36 + (-1)*self_BFI_51 + 
(-1)*self_BFI_11 + (-1)*self_BFI_26 + 1*self_BFI_41 + 1*self_BFI_56
'
```
Model 2 has a negative wording factor to capture the residuals variance of reversely worded items.
```{r}
Model2_E <- '
Sociability  =~ self_BFI_1 + self_BFI_16 + self_BFI_31 + self_BFI_46
Assertiveness  =~ self_BFI_6 + self_BFI_21 + self_BFI_36 + self_BFI_51
Energy_Level  =~ self_BFI_11 + self_BFI_26 + self_BFI_41 + self_BFI_56

#Extroversion =~ Sociability + Assertiveness + Energy_Level

Sociability ~~ 0*Reverse
Energy_Level ~~ 0*Reverse
Assertiveness ~~ 0*Reverse

Reverse =~ self_BFI_16 + self_BFI_31 + self_BFI_36 + self_BFI_51 + self_BFI_11 + self_BFI_26 
'
```
Model 3 has a positive wording factor to capture the residuals variance of positively worded items. The alternatice syntax is used for the Expanded format to constrain the variance to be positive
```{r}
Model3_E <- '
Sociability  =~ self_BFI_1 + self_BFI_16 + self_BFI_31 + self_BFI_46
Assertiveness  =~ self_BFI_6 + self_BFI_21 + self_BFI_36 + self_BFI_51
Energy_Level  =~ self_BFI_11 + self_BFI_26 + self_BFI_41 + self_BFI_56

#Extroversion =~ Sociability + Assertiveness + Energy_Level

Sociability ~~ 0*Positive
Energy_Level ~~ 0*Positive
Assertiveness ~~ 0*Positive

Positive =~ self_BFI_1 + self_BFI_46 + self_BFI_6 + self_BFI_21 + self_BFI_41 + self_BFI_56
'

# The Expanded data do not fit Model 3. To converge the mode, we restricted the residual 
# of item 56 to positive 
Model3_E_alternative <- '
Sociability  =~ self_BFI_1 + self_BFI_16 + self_BFI_31 + self_BFI_46
Assertiveness  =~ self_BFI_6 + self_BFI_21 + self_BFI_36 + self_BFI_51
Energy_Level  =~ self_BFI_11 + self_BFI_26 + self_BFI_41 + self_BFI_56

#Extroversion =~ Sociability + Assertiveness + Energy_Level

Sociability ~~ 0*Positive
Energy_Level ~~ 0*Positive
Assertiveness ~~ 0*Positive

self_BFI_56~~ a* self_BFI_56
a>0

Positive =~ self_BFI_1 + self_BFI_46 + self_BFI_6 + self_BFI_21 + self_BFI_41 + self_BFI_56
'
```
#### Constructing the Models 
```{r}
Run_Model0_Likert_E <- sem(Model0_E,dat = likert_data, missing = "ML")
Run_Model1_Likert_E <- sem(Model1_E,dat = likert_data, missing = "ML")
Run_Model2_Likert_E <- sem(Model2_E,dat = likert_data,missing = "ML")
Run_Model3_Likert_E <- sem(Model3_E,dat = likert_data,missing = "ML")

Run_Model0_Expanded_E <- sem(Model0_E,dat = expanded_data, missing = "ML")
Run_Model1_Expanded_E <- sem(Model1_E,dat = expanded_data, missing = "ML")
Run_Model2_Expanded_E <- sem(Model2_E,dat = expanded_data, missing = "ML")
Run_Model3_Expanded_E <- sem(Model3_E_alternative,dat = expanded_data, missing = "ML") # the modified Model 3
```

#### Descriptive Statistics for Model 0  
To obtain the coefficient omega
```{r}
reliability(Run_Model0_Likert_E,return.total=TRUE)
reliability(Run_Model0_Expanded_E,return.total=TRUE)
```
To obtain the average item mean
```{r}
mean(colMeans(likert_data[,c(1 , 16 , 31 , 46, 6 , 21 , 36 , 51, 11 , 26 , 41 , 56)],na.rm=TRUE))
mean(colMeans(expanded_data[,c(1 , 16 , 31 , 46, 6 , 21 , 36 , 51, 11 , 26 , 41 , 56)],na.rm=TRUE))
```
To obtain the average item variance
```{r}
mean(sapply(likert_data[,c(1 , 16 , 31 , 46, 6 , 21 , 36 , 51, 11 , 26 , 41 , 56)], var, na.rm=TRUE))
mean(sapply(expanded_data[,c(1 , 16 , 31 , 46, 6 , 21 , 36 , 51, 11 , 26 , 41 , 56)], var, na.rm=TRUE))
```

#### Model Summary
##### Likert Format
```{r}
summary(Run_Model0_Likert_E, standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model1_Likert_E,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model2_Likert_E,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model3_Likert_E,standardized = TRUE,fit.measures = TRUE)
```
##### Expanded Format
```{r}
summary(Run_Model0_Expanded_E, standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model1_Expanded_E,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model2_Expanded_E,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model3_Expanded_E,standardized = TRUE,fit.measures = TRUE)
```

#### Model Fit indices 
```{r}
summary(compareFit(Run_Model0_Likert_E,Run_Model1_Likert_E,Run_Model2_Likert_E,Run_Model3_Likert_E,nested = FALSE))
```

```{r}
summary(compareFit(Run_Model0_Expanded_E,Run_Model1_Expanded_E,Run_Model2_Expanded_E,Run_Model3_Expanded_E,nested = FALSE))
```


### Agreeableness
#### Model Syntax
Model 0 has no method factor.

```{r}
Model0_A <- '
Compassion  =~ self_BFI_2 + self_BFI_17 + self_BFI_32 + self_BFI_47
Respectfulness  =~ self_BFI_7 + self_BFI_22 + self_BFI_37 + self_BFI_52
Trust  =~ self_BFI_12 + self_BFI_27 + self_BFI_42 + self_BFI_57

#Agreeableness =~ Compassion + Respectfulness + Trust
'
```
Model 1 has an Acquiescence factor to capture the residuals variance of all items. 
```{r}
Model1_A <- '
Compassion  =~ self_BFI_2 + self_BFI_17 + self_BFI_32 + self_BFI_47
Respectfulness  =~ self_BFI_7 + self_BFI_22 + self_BFI_37 + self_BFI_52
Trust  =~ self_BFI_12 + self_BFI_27 + self_BFI_42 + self_BFI_57

#Agreeableness =~ Compassion + Respectfulness + Trust

Compassion ~~ 0*Acquiescence
Respectfulness ~~ 0*Acquiescence
Trust ~~ 0*Acquiescence

Acquiescence =~ 1*self_BFI_2 + (-1)*self_BFI_17 + 1*self_BFI_32 + (-1)*self_BFI_47 + 
1*self_BFI_7 + (-1)*self_BFI_22 + (-1)*self_BFI_37 + 1*self_BFI_52 + 
(-1)*self_BFI_12 + 1*self_BFI_27 + (-1)*self_BFI_42 + 1*self_BFI_57
'
```
Model 2 has a negative wording factor to capture the residuals variance of reversely worded items.
```{r}
Model2_A <- '
Compassion  =~ self_BFI_2 + self_BFI_17 + self_BFI_32 + self_BFI_47
Respectfulness  =~ self_BFI_7 + self_BFI_22 + self_BFI_37 + self_BFI_52
Trust  =~ self_BFI_12 + self_BFI_27 + self_BFI_42 + self_BFI_57

#Agreeableness =~ Compassion + Respectfulness + Trust

Compassion ~~ 0*Reverse
Respectfulness ~~ 0*Reverse
Trust ~~ 0*Reverse

Reverse =~ self_BFI_17 + self_BFI_47 + self_BFI_22 + self_BFI_37 + self_BFI_12 + self_BFI_42
'
```
Model 3 has a positive wording factor to capture the residuals variance of positively worded items. The alternatice syntax is used for the Expanded format to constrain the variance to be positive
```{r}
Model3_A <- '
Compassion  =~ self_BFI_2 + self_BFI_17 + self_BFI_32 + self_BFI_47
Respectfulness  =~ self_BFI_7 + self_BFI_22 + self_BFI_37 + self_BFI_52
Trust  =~ self_BFI_12 + self_BFI_27 + self_BFI_42 + self_BFI_57

#Agreeableness =~ Compassion + Respectfulness + Trust

Compassion ~~ 0*Positive
Respectfulness ~~ 0*Positive
Trust ~~ 0*Positive

Positive =~ self_BFI_2 + self_BFI_32 + self_BFI_7 + self_BFI_52 + self_BFI_27 + self_BFI_57
'
```
#### Constructing the Models 
```{r}
Run_Model0_Likert_A <- sem(Model0_A,dat = likert_data, missing = "ML")
Run_Model1_Likert_A <- sem(Model1_A,dat = likert_data, missing = "ML")
Run_Model2_Likert_A <- sem(Model2_A,dat = likert_data,missing = "ML") # This Model does not converge 
Run_Model3_Likert_A <- sem(Model3_A,dat = likert_data,missing = "ML")

Run_Model0_Expanded_A <- sem(Model0_A,dat = expanded_data, missing = "ML")
Run_Model1_Expanded_A <- sem(Model1_A,dat = expanded_data, missing = "ML")
Run_Model2_Expanded_A <- sem(Model2_A,dat = expanded_data, missing = "ML") # This Model does not converge 
Run_Model3_Expanded_A <- sem(Model3_A,dat = expanded_data, missing = "ML") 
```

#### Descriptive Statistics for Model 0  
To obtain the coefficient omega
```{r}
reliability(Run_Model0_Likert_A,return.total=TRUE)
reliability(Run_Model0_Expanded_A,return.total=TRUE)
```
To obtain the average item mean
```{r}
mean(colMeans(likert_data[,c(2, 7, 12, 17, 22, 27, 32, 37, 42, 47, 52, 57)],na.rm=TRUE))
mean(colMeans(expanded_data[,c(2, 7, 12, 17, 22, 27, 32, 37, 42, 47, 52, 57)],na.rm=TRUE))
```
To obtain the average item variance
```{r}
mean(sapply(likert_data[,c(2, 7, 12, 17, 22, 27, 32, 37, 42, 47, 52, 57)], var, na.rm=TRUE))
mean(sapply(expanded_data[,c(2, 7, 12, 17, 22, 27, 32, 37, 42, 47, 52, 57)], var, na.rm=TRUE))
```
#### Model Summary
##### Likert Format
```{r}
summary(Run_Model0_Likert_A, standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model1_Likert_A,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model2_Likert_A,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model3_Likert_A,standardized = TRUE,fit.measures = TRUE)
```
##### Expanded Format
```{r}
summary(Run_Model0_Expanded_A, standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model1_Expanded_A,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model2_Expanded_A,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model3_Expanded_A,standardized = TRUE,fit.measures = TRUE)
```
#### Model Fit indices 
```{r}
summary(compareFit(Run_Model0_Likert_A,Run_Model1_Likert_A,Run_Model2_Likert_A,Run_Model3_Likert_A,nested = FALSE))
```

```{r}
summary(compareFit(Run_Model0_Expanded_A,Run_Model1_Expanded_A,Run_Model2_Expanded_A,Run_Model3_Expanded_A,nested = FALSE))
```

### Consciousness
#### Model Syntax
Model 0 has no method factor.

```{r}
Model0_C <- '
Organization  =~ self_BFI_3 + self_BFI_18 + self_BFI_33 + self_BFI_48
Productiveness  =~ self_BFI_8 + self_BFI_23 + self_BFI_38 + self_BFI_53
Responsibility  =~ self_BFI_13 + self_BFI_28 + self_BFI_43 + self_BFI_58

#Consciousness =~ Organization + Productiveness + Responsibility
'
```
Model 1 has an Acquiescence factor to capture the residuals variance of all items. 
```{r}
Model1_C <- '
Organization  =~ self_BFI_3 + self_BFI_18 + self_BFI_33 + self_BFI_48
Productiveness  =~ self_BFI_8 + self_BFI_23 + self_BFI_38 + self_BFI_53
Responsibility  =~ self_BFI_13 + self_BFI_28 + self_BFI_43 + self_BFI_58

#Consciousness =~ Organization + Productiveness + Responsibility

Organization ~~ 0*Acquiescence
Productiveness ~~ 0*Acquiescence
Responsibility ~~ 0*Acquiescence

Acquiescence =~ (-1)*self_BFI_3 + 1*self_BFI_18 + 1*self_BFI_33 + (-1)*self_BFI_48 + 
(-1)*self_BFI_8 + (-1)*self_BFI_23 + 1*self_BFI_38 + 1*self_BFI_53 + 
1*self_BFI_13 + (-1)*self_BFI_28 + 1*self_BFI_43 + (-1)*self_BFI_58
'
```
Model 2 has a negative wording factor to capture the residuals variance of reversely worded items.
```{r}
Model2_C <- '
Organization  =~ self_BFI_3 + self_BFI_18 + self_BFI_33 + self_BFI_48
Productiveness  =~ self_BFI_8 + self_BFI_23 + self_BFI_38 + self_BFI_53
Responsibility  =~ self_BFI_13 + self_BFI_28 + self_BFI_43 + self_BFI_58

#Consciousness =~ Organization + Productiveness + Responsibility

Organization ~~ 0*Reverse
Productiveness ~~ 0*Reverse
Responsibility ~~ 0*Reverse

Reverse =~ self_BFI_3 + self_BFI_48 + self_BFI_8 + self_BFI_23  + self_BFI_28 + self_BFI_58
'
```
Model 3 has a positive wording factor to capture the residuals variance of positively worded items. The alternatice syntax is used for the Expanded format to constrain the variance to be positive
```{r}
Model3_C <- '
Organization  =~ self_BFI_3 + self_BFI_18 + self_BFI_33 + self_BFI_48
Productiveness  =~ self_BFI_8 + self_BFI_23 + self_BFI_38 + self_BFI_53
Responsibility  =~ self_BFI_13 + self_BFI_28 + self_BFI_43 + self_BFI_58

#Consciousness =~ Organization + Productiveness + Responsibility

Organization ~~ 0*Positive
Productiveness ~~ 0*Positive
Responsibility ~~ 0*Positive

Positive =~ self_BFI_18 + self_BFI_33 + self_BFI_38 + self_BFI_53 + self_BFI_13 + self_BFI_43 
'
```

#### Constructing the Models 
```{r}
Run_Model0_Likert_C <- sem(Model0_C,dat = likert_data, missing = "ML")
Run_Model1_Likert_C <- sem(Model1_C,dat = likert_data, missing = "ML")
Run_Model2_Likert_C <- sem(Model2_C,dat = likert_data,missing = "ML") 
Run_Model3_Likert_C <- sem(Model3_C,dat = likert_data,missing = "ML")

Run_Model0_Expanded_C <- sem(Model0_C,dat = expanded_data, missing = "ML")
Run_Model1_Expanded_C <- sem(Model1_C,dat = expanded_data, missing = "ML")
Run_Model2_Expanded_C <- sem(Model2_C,dat = expanded_data, missing = "ML") # Negative Estimate or variances
Run_Model3_Expanded_C <- sem(Model3_C,dat = expanded_data, missing = "ML") 
```

#### Descriptive Statistics for Model 0  
To obtain the coefficient omega
```{r}
reliability(Run_Model0_Likert_C,return.total=TRUE)
reliability(Run_Model0_Expanded_C,return.total=TRUE)
```
To obtain the average item mean
```{r}
mean(colMeans(likert_data[,c(3, 8, 13, 18, 23, 28, 33, 38, 43, 48, 53, 58)],na.rm=TRUE))
mean(colMeans(expanded_data[,c(3, 8, 13, 18, 23, 28, 33, 38, 43, 48, 53, 58)],na.rm=TRUE))
```
To obtain the average item variance
```{r}
mean(sapply(likert_data[,c(3, 8, 13, 18, 23, 28, 33, 38, 43, 48, 53, 58)], var, na.rm=TRUE))
mean(sapply(expanded_data[,c(3, 8, 13, 18, 23, 28, 33, 38, 43, 48, 53, 58)], var, na.rm=TRUE))
```

#### Model Summary
##### Likert Format
```{r}
summary(Run_Model0_Likert_C, standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model1_Likert_C,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model2_Likert_C,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model3_Likert_C,standardized = TRUE,fit.measures = TRUE)
```
##### Expanded Format
```{r}
summary(Run_Model0_Expanded_C, standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model1_Expanded_C,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model2_Expanded_C,standardized = TRUE,fit.measures = TRUE)
```
```{r}
summary(Run_Model3_Expanded_C,standardized = TRUE,fit.measures = TRUE)
```
#### Model Fit indices 
```{r}
summary(compareFit(Run_Model0_Likert_C,Run_Model1_Likert_C,Run_Model2_Likert_C,Run_Model3_Likert_C,nested = FALSE))
```

```{r}
summary(compareFit(Run_Model0_Expanded_C,Run_Model1_Expanded_C,Run_Model2_Expanded_C,Run_Model3_Expanded_C,nested = FALSE))
```